---
title: "R Notebook, analysis of brain samples, different sections"
author: "Constantinos Yeles"
date: "`r format(Sys.time(), '%a_%b_%d_%Y')`"
output:
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---
## load libraries
```{r load libraries}
suppressPackageStartupMessages(library('tidyverse'))
suppressPackageStartupMessages(library('data.table'))
```
## add date
```{r date_of_analysis}
date <- format(Sys.time(), "%d_%b_%Y")
```
## create the dir for the analysis
```{r}
my_exp <- "brain_diff_sect"
dir.create(str_glue("/home/0/SPAR_data_analysis/{my_exp}_analysis"))
dat_path <- str_glue("{my_exp}_analysis")
```

## import data and make dataframe
```{r create_the_df}
path <- "/home/0/my_data/SPAR_res"
smallRNA_files <- dir(path, full.names = TRUE,
                      pattern = "smRNA_gene_expression.xls",
                      recursive = TRUE)
# load the list of files in one table -----
DT <- rbindlist(sapply(smallRNA_files,fread,
                       simplify=FALSE,
                       #select = c(1,2,4,6),
                       verbose=getOption("datatable.verbose", TRUE)),
                use.names= TRUE,idcol="file")  %>%
  rename(smallRNA = "#Gene")

# sep "smallRNA" column to multiple
# separate("smallRNA",c("chr","start","end","strand","smallRNA","DQ"), sep = ":") %>%
#  as_tibble()

# make matrix for DE analysis  -----
dt <- DT %>% group_by(file) %>% select(-RPM) %>%
  spread(key = "file", value = "ReadCount" )
# clean the colnames
names(dt)[3:length(names(dt))] <- names(dt)[3:length(names(dt))] %>%
  str_remove(".+/SPAR_res/") %>%
  str_remove(".trimmed_.+") %>%
  str_replace("^?","S")

# load the targets file ----
targ_path <- dir(path = path, 
                 pattern = "targets.+.csv",
                 full.names = T,
                 recursive = TRUE)

targets_file <- fread(targ_path) %>%
  as_tibble() %>% 
  mutate(Sample_id = str_replace(Sample_id,"^?","S"))

targets_file <- targets_file %>% 
  filter(Tissue == "Brain",
         Dev_stage == "Adult") %>% 
  select(samples, Sample_id, Date_of_seq, Total_region,
         Reads, Sex, Age, Donors) %>% 
  mutate_if(is.character, as_factor)

target <- match(names(dt),targets_file$Sample_id, nomatch = 0)
tardt <- match(targets_file$Sample_id, names(dt), nomatch = 0)
#dt1 <- dt
names(dt)[tardt] <- as.character(targets_file$samples[target])

#tibble(names(dt)[3:length(names(dt))],targets_file$Sample_id)
#tibble(names(dt1), names(dt))

stopifnot(identical(targets_file$Sample_id,names(dt)[3:length(names(dt))]))

# write the results  -----
dt %>% write_tsv(str_glue("{dat_path}/raw_reads_{date}_{my_exp}_SPAR.txt"))
```

## Perform various Filtering Methods: EdgeR, NOIseq
```{r}
suppressPackageStartupMessages(library('edgeR'))
suppressPackageStartupMessages(library('NOIseq'))# there is a problem loading it 
# I load it manually 
suppressPackageStartupMessages(library('rafalib'))

#subset the dataset
dt <- dt[c(1,2, tardt)]
targets_file <- targets_file %>% filter(Sample_id %in% colnames(dt))
# make the matrix  -----
mat <- dt %>%
  select(-GeneClass) %>%
  column_to_rownames("smallRNA") %>%
  as.matrix()

# filtering with NOISEq  -----
noiL <- as.list(1L:3)
names(noiL) <- c("cpm", "Wilcoxon", "Prop")
NOISeq_filt <- lapply(noiL, function(x) NOISeq::filtered.data(mat, 
        factor = targets_file$Total_region,
        norm = FALSE, 
        method = x, cv.cutoff=100, cpm=1))

# create the design matrix  -----
design <- model.matrix(~targets_file$Total_region)
colnames(design) <- gsub("targets_file\\$Total_region","",colnames(design))
rownames(design) <- targets_file$Total_region

# create the DGElist ----
targets_file <- targets_file %>% dplyr::rename(group = "Total_region")
dgl <- DGEList(counts = mat,
               samples = targets_file,
               lib.size = colSums(mat),
               norm.factors = rep(1,ncol(mat)))

# filterwith EdgeR ----
keep.exprs <- filterByExpr.DGEList(dgl, design = design)
dgl_filt <- dgl[keep.exprs,,keep.lib.sizes=FALSE]

str_glue( "Starting features: {nrow(dgl)}\nedgeR filter, remaining features: {nrow(dgl_filt)}\nCPM approach, remaining features: {nrow(NOISeq_filt[[1]])}\nWilcoxon filter, remaining features: {nrow(NOISeq_filt[[2]])}\nProportion approach filter, remaining features: {nrow(NOISeq_filt[[3]])}")

# if we want to keep noiseq filter instead of edger but use edger object ----
filte_Nois <- dgl[rownames(dgl$counts) %in% rownames(NOISeq_filt[[3]]),,keep.lib.sizes=FALSE]

dgl_filt <- DGEList(counts = NOISeq_filt[[3]],
               samples = targets_file,
               lib.size = colSums(NOISeq_filt[[3]]),
               norm.factors = rep(1,ncol(NOISeq_filt[[3]])))
# finally give colours to samples ----
suppressPackageStartupMessages(library('wesanderson'))
pal1 <- wes_palette(length(levels(as_factor(targets_file$group))),
                    name = "Darjeeling1", type = "continuous") %>%
  enframe(name = NULL) %>%
  mutate(
    group = as_factor(levels(targets_file$group)))

# add color to the DGEList 
dgl_filt$colours <- inner_join(dgl_filt$samples,pal1,by= "group")$value


# plot RLE and other functions from EDASeq ----
suppressPackageStartupMessages(library('EDASeq'))
EDASeq::plotRLE(dgl_filt$counts, col=dgl_filt$colours, 
                outline=FALSE, las=3, ylab="Relative Log Expression", 
                cex.axis=1, cex.lab=1)
EDASeq::plotPCA(dgl_filt$counts, 
                col=dgl_filt$colours, cex=1, cex.axis=1, cex.lab=1)

dataEDA <- EDASeq::newSeqExpressionSet(counts = dgl_filt$counts, 
  phenoData = as.data.frame(dgl_filt$samples, row.names = rownames(dgl_filt$samples)))
EDASeq::meanVarPlot(dataEDA, log=TRUE, ylim=c(0,16))
EDASeq::biasPlot(dataEDA, "gc", log=TRUE, ylim=c(1,5))
```
## histogram
```{r}
AveLogCpm_Raw_Data <- aveLogCPM(dgl)
AveLogCpm_Filtered_Data <-aveLogCPM(dgl_filt)

par(mfrow=c(1,2))

pdf(str_glue("{dat_path}/histogram_plot_{date}_{my_exp}.pdf"))
hist(AveLogCpm_Raw_Data)
hist(AveLogCpm_Filtered_Data)
dev.off()
```

## proceed with the normalization
```{r}
# not normalized
cpm_unorm <- cpm(dgl_filt,log=F,prior.count=5)
lcpm_unorm <- cpm(dgl_filt,log=TRUE,prior.count=5)
# EDA norm ----
dataOffset <- betweenLaneNormalization(dataEDA,
                                       which="upper",offset=TRUE)

EDASeq::meanVarPlot(dataOffset, log=TRUE, ylim=c(0,16))
# edgeR  ----
norm_method <- list("none" = "none", 
                  "TMM" = "TMM",
                  "TMMwsp" = "TMMwsp",
                  "RLE" = "RLE" )
edger_norm <- lapply(norm_method, function(x){
  calcNormFactors(dgl_filt, method = x)
})

#x1_TMM <- calcNormFactors(dgl_filt, method = "TMM")
#cpm_TMM <- cpm(x1_TMM,normalized.lib.sizes = TRUE)
#lcpm_TMM <- cpm(x1_TMM,normalized.lib.sizes = TRUE, log=TRUE, prior.count=5)

# voom with quality weights ----
voomQW_norm <- mapply(function(x, i)
  voomWithQualityWeights(x, design = design, plot = TRUE, 
                         normalize.method = i), edger_norm[1:3], 
  c("quantile", rep("none",2)), SIMPLIFY = FALSE 
)
# create list of normalized data ------

norm_list <- list("Not normalized" = cpm(edger_norm[[1]],normalized.lib.sizes = TRUE), 
                  "EDA_upperQ" = dataOffset, 
                  "TMM" = cpm(edger_norm[[2]],normalized.lib.sizes = TRUE),
                  "TMMwsp" = cpm(edger_norm[[3]],normalized.lib.sizes = TRUE), 
                  "RLE"  = cpm(edger_norm[[4]],normalized.lib.sizes = TRUE),
                  "voom_Quantile" = 2^voomQW_norm[[1]]$E,
                  "voom_TMM" = 2^voomQW_norm[[2]]$E,
                  "voom_TMMwsp" = 2^voomQW_norm[[3]]$E)

pdf(str_glue("{dat_path}/RLE_plots_{date}_{my_exp}.pdf"))

sapply(seq_along(norm_list) ,function(x, n, i){ 
       EDASeq::plotRLE(x[[i]], col=dgl_filt$colours,
                       outline=FALSE, las=3, 
                       ylab="Relative Log Expression", 
                       cex.axis=1, cex.lab=1, main = str_glue("{n[[i]]}"));
  legend("topright",
       legend= levels(as_factor(dgl_filt$samples$group)),
       fill = levels(as_factor(dgl_filt$colours)),
       bty="n",
       cex = 0.5, inset = c(.01,.01) )}, x = norm_list, n = names(norm_list))
dev.off()


#cpm_TMM <- cpm(x1_TMM,normalized.lib.sizes = TRUE, prior.count=5)

#EDASeq::plotRLE(cpm_TMM, col=dgl_filt$colours, 
#                outline=FALSE, las=3, ylab="Relative Log Expression", 
#                cex.axis=1, cex.lab=1, main = "TMM")

#write_rds(x1_TMM,str_glue("x1_TMMwzp_{experiment}_{date}.rds"))
#x1_voomTMM <- read_rds("~/Documents/new_analysis_workflow/results_R/results_with_DBs/3_IPP_PIWIL1_all_DBs/x1_voomTMM.rds")
```

## make h-clustering
```{r Hierarchical clustering}
norm_l <- list(
               "unorm" = cpm(edger_norm[[1]],normalized.lib.sizes = TRUE, log=TRUE, prior.count=5),
               "TMM" = cpm(edger_norm[[2]],normalized.lib.sizes = TRUE, log=TRUE, prior.count=5),
               "TMMwsp" = cpm(edger_norm[[3]],normalized.lib.sizes = TRUE, log=TRUE, prior.count=5),
               "voomQ_QW" = voomQW_norm[[1]]$E,
               "voomTMM_QW" = voomQW_norm[[2]]$E,
               "voomTMM_QW" = voomQW_norm[[3]]$E)

hc_methods <- c("ward.D2",
                "single",
                "complete",
                "average",
                #"median", "centroid",
                "mcquitty")
l_d <- lapply(norm_l,function(x)dist(t(x)))

l_hc <- sapply(hc_methods, function(x) lapply(l_d, function(y) hclust(y,method = x)))

names(l_hc) <- rep(rownames(l_hc),times = ncol(l_hc))

pdf(str_glue("{dat_path}/hierarchic_clust_{date}_{my_exp}.pdf"))
for (i in seq_along(l_hc)) {
  myplclust(l_hc[[i]],
            lab.col = dgl_filt$colours, #x1$colours,pal1$value
            labels = dgl_filt$samples$group,#pal1$group
       xlab = str_glue("{names(l_hc[i])}"),
       main = str_glue("{enframe(l_hc[[i]])$value[7]} - {enframe(l_hc[[i]])$value[5]} - group"))
  legend("topright",
       legend = levels(as_factor(dgl_filt$samples$group)),
       fill = levels(as_factor(dgl_filt$colours)), #x1$colours  
       bty="n",
       cex = 0.9)
}

#sapply(l_hc,function(x)myplclust(x,lab.col = x1$colours,labels = x1_TMM$samples$group,
#       xlab = str_glue("{names(x)}"), main = str_glue("{enframe(x)$value[7]} - {enframe(x)$value[5]} - region")))

dev.off()

#d <- dist(t(lcpm_TMM))
#par(mar=c(7,3,2,2))
#myplclust(hc,lab.col = x1$colours,labels = x1_TMM$samples$group,
#          xlab = "",main="euclidean-ward.D2 - region")
#myplclust(hc,lab.col = x1$colours,labels = targets_file$Years,
#          xlab = "",main="euclidean-ward.D2 - age")
#myplclust(hc,lab.col = x1$colours,labels = targets_file$data_sequenziamento,
#          xlab = "",main="euclidean-ward.D2 - year of run")

#legend("topright",legend = levels(group),
#       fill = x1$colours,
#       bty="n",
#       cex = 0.6)

```

## remove unwanted variation
```{r}
suppressPackageStartupMessages(library('sva'))
# full model
mod <-  model.matrix(~dgl_filt$samples$group, data = dgl_filt$samples)
colnames(mod) <- gsub("dgl_filt\\$samples\\$group","",colnames(mod))

# null model
mod0 <- cbind(mod[,1])

# estimate batch and other artifacts
svseq <- svaseq(dgl_filt$counts, mod, mod0, n.sv = 1)
plot(svseq$sv, pch=19, col="blue")

n.sv <- num.sv(voomQW_norm[[1]]$E, mod, method = "leek")
svobj <- sva(voomQW_norm[[1]]$E, mod, mod0,n.sv = n.sv)



modSv <- cbind(mod,svobj$sv)
fit = lmFit(voomQW_norm[[1]],modSv)
```


The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
