---
title: "R Notebook, analysis of brain samples, GSE63501"
author: "Constantinos Yeles"
todate: "`r format(Sys.time(), '%a_%b_%d_%Y')`"
output:
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---
## load libraries
```{r load libraries}
suppressPackageStartupMessages({
  library('tidyverse')
  library('data.table')
})
```
## add todate
```{r todate_of_analysis}
todate <- format(Sys.time(), "%d_%b_%Y")
```
## create the dir for the analysis
```{r}
my_exp <- "GSE63501"
dir.create(str_glue("/home/0/SPAR_data_analysis/{my_exp}_analysis"))
dat_path <- str_glue("{my_exp}_analysis")
```

## import data and make dataframe
```{r create_the_df}
path <- "/home/0/geobrain/GSE63501_16_CTRL_AD"
smallRNA_files <- dir(path, full.names = TRUE,
                      pattern = "smRNA_gene_expression.xls",
                      recursive = TRUE)
# load the list of files in one table -----
DT <- rbindlist(sapply(smallRNA_files,fread,
                       simplify=FALSE,
                       #select = c(1,2,4,6),
                       verbose=getOption("datatable.verbose", TRUE)),
                use.names= TRUE,idcol="file")  %>%
  rename(smallRNA = "#Gene")

# sep "smallRNA" column to multiple
# separate("smallRNA",c("chr","start","end","strand","smallRNA","DQ"), sep = ":") %>%
#  as_tibble()

# make matrix for DE analysis  -----
dt <- DT %>% group_by(file) %>% select(-RPM) %>%
  spread(key = "file", value = "ReadCount" )

# clean the colnames
names(dt)[3:length(names(dt))] <- names(dt)[3:length(names(dt))] %>%
  str_remove(".+/GSE63501_16_CTRL_AD/") %>%
  str_remove(".trimmed_.+") 

# load the targets file ----
targ_path <- dir(path = path, 
                 pattern = "Sra",
                 full.names = T,
                 recursive = TRUE)

targets_file <- fread(targ_path) %>% 
  as_tibble() %>% 
  dplyr::select(Run, BioSample, classification, Instrument) %>% 
  mutate_if(is.character, as_factor) 

target <- match(names(dt),targets_file$Run, nomatch = 0)
tardt <- match(targets_file$Run, names(dt), nomatch = 0)
#dt1 <- dt
names(dt)[tardt] <- as.character(targets_file$samples[target])

stopifnot(identical(targets_file$Run,names(dt)[3:length(names(dt))]))

# write the results  -----
dt %>% write_tsv(str_glue("{dat_path}/raw_reads_{todate}_{my_exp}_SPAR.txt"))
```

## Perform various Filtering Methods: EdgeR, NOIseq
```{r}
# load dataset ----
DT <- fread(list.files(path = dat_path, pattern = "raw", full.names = TRUE ))
suppressPackageStartupMessages({library('edgeR')
  library('NOIseq')})# there is a problem loading it 
# I load it manually 
suppressPackageStartupMessages(library('rafalib'))
stopifnot(identical(as.character(targets_file$Run),names(dt)[3:length(names(dt))]))
# exploratory analysis with NOISeq -----
mybiotypes <-  dt$GeneClass
names(mybiotypes) <-  dt$smallRNA

# make the matrix  
mat <- dt %>%
  select(-GeneClass) %>%
  column_to_rownames("smallRNA") %>%
  as.matrix()
# make clearer names
colnames(mat) <- str_glue("{targets_file$classification}_{colnames(mat)}") %>% 
  str_replace("Control","Ctl")
targets_file$Run <- colnames(mat)

# NOISeq dataset
mydata <- readData(data = mat, factors = as.data.frame(targets_file), biotype = mybiotypes)

mybiodetection <- dat(mydata, k = 0, type = "biodetection", factor = NULL)
par(mfrow = c(1, 2))
explo.plot(mybiodetection, samples = c(1,3), toplot = "piRNA",plottype = "comparison")
mycountsbio <- dat(mydata, factor = NULL, type = "countsbio")
par(mfrow = c(1,1 ))

pdf(str_glue("{dat_path}/biodetection_expression_{todate}_{my_exp}.pdf"))
sapply(seq(ncol(mat)) ,function(x){
explo.plot(mycountsbio, toplot = 2, samples = x, plottype = "boxplot")
})
dev.off()

# remove noisy samples ----
mat <- mat[,!colnames(mat) %in% c("Ctl_SRR1658345","AD_SRR1658353")]
targets_file <- targets_file %>% filter(!Run %in% c("Ctl_SRR1658345","AD_SRR1658353"))
# filtering with NOISEq  -----
noiL <- as.list(1L:3)
names(noiL) <- c("cpm", "Wilcoxon", "Prop")
NOISeq_filt <- lapply(noiL, function(x) NOISeq::filtered.data(mat, 
        factor = targets_file$classification,
        norm = FALSE, 
        method = x, cv.cutoff=100, cpm=1))

# create the design matrix  -----
targets_file <- targets_file %>% dplyr::rename(group = "classification")
design <- model.matrix(~targets_file$group)
colnames(design) <- gsub("targets_file\\$group","",colnames(design))
rownames(design) <- targets_file$group

# create the DGElist ----
dgl <- DGEList(counts = mat,
               samples = targets_file,
               lib.size = colSums(mat),
               norm.factors = rep(1,ncol(mat)))

# filterwith EdgeR ----
keep.exprs <- filterByExpr.DGEList(dgl, design = design)
dgl_filt <- dgl[keep.exprs,,keep.lib.sizes=FALSE]

dim(dgl);dim(dgl_filt)
str_glue( "Starting features: {nrow(dgl)}\nedgeR filter, remaining features: {nrow(dgl_filt)}\nCPM approach, remaining features: {nrow(NOISeq_filt[[1]])}\nWilcoxon filter, remaining features: {nrow(NOISeq_filt[[2]])}\nProportion approach filter, remaining features: {nrow(NOISeq_filt[[3]])}")

# if we want to keep noiseq filter instead of edger but use edger object ----
filte_Nois <- dgl[rownames(dgl$counts) %in% rownames(NOISeq_filt[[i]]),,keep.lib.sizes=FALSE]

dgl_filt <- DGEList(counts = NOISeq_filt[[i]],
               samples = targets_file,
               lib.size = colSums(NOISeq_filt[[i]]),
               norm.factors = rep(1,ncol(NOISeq_filt[[i]])))

# finally give colours to samples ----
suppressPackageStartupMessages(library('wesanderson'))
pal1 <- wes_palette(length(levels(as_factor(targets_file$group))),
                    name = "Darjeeling1", type = "discrete") %>%
  enframe(name = NULL) %>%
  mutate(
    group = as_factor(levels(targets_file$group)))

# add color to the DGEList 
dgl_filt$colours <- inner_join(dgl_filt$samples,pal1,by= "group")$value

# plot RLE and other functions from EDASeq ----
suppressPackageStartupMessages(library('EDASeq'))
par(mar=c(8,4,4,2))
EDASeq::plotRLE(dgl_filt$counts, col=dgl_filt$colours, 
                outline=FALSE,las = 2, ylab="Relative Log Expression", 
                cex.axis=1, cex.lab=1, )
EDASeq::plotPCA(dgl_filt$counts, 
                col = dgl_filt$colours, cex=1, cex.axis=1, cex.lab=1)

dataEDA <- EDASeq::newSeqExpressionSet(counts = dgl_filt$counts, 
  phenoData = as.data.frame(dgl_filt$samples, row.names = rownames(dgl_filt$samples)))
EDASeq::meanVarPlot(dataEDA, log=TRUE, ylim=c(0,16))
EDASeq::biasPlot(dataEDA, "gc", log=TRUE, ylim=c(1,5))
```
## histogram
```{r}
AveLogCpm_Raw_Data <- aveLogCPM(dgl)
AveLogCpm_Filtered_Data <-aveLogCPM(dgl_filt)

par(mfrow=c(1,2))

pdf(str_glue("{dat_path}/histogram_plot_{todate}_{my_exp}.pdf"))
hist(AveLogCpm_Raw_Data)
hist(AveLogCpm_Filtered_Data)
dev.off()
```

## proceed with the normalization
```{r}
# not normalized
cpm_unorm <- cpm(dgl_filt,log=F,prior.count=5)
lcpm_unorm <- cpm(dgl_filt,log=TRUE,prior.count=5)
# EDA norm ----
dataOffset <- betweenLaneNormalization(dataEDA,
                                       which="upper",offset=TRUE)

EDASeq::meanVarPlot(dataOffset, log=TRUE, ylim=c(0,16))
# edgeR  ----
norm_method <- list("none" = "none", 
                  "TMM" = "TMM",
                  "TMMwsp" = "TMMwsp",
                  "RLE" = "RLE" )
edger_norm <- lapply(norm_method, function(x){
  calcNormFactors(dgl_filt, method = x)
})

#x1_TMM <- calcNormFactors(dgl_filt, method = "TMM")
#cpm_TMM <- cpm(x1_TMM,normalized.lib.sizes = TRUE)
#lcpm_TMM <- cpm(x1_TMM,normalized.lib.sizes = TRUE, log=TRUE, prior.count=5)

# voom with quality weights ----
voomQW_norm <- mapply(function(x, i)
  voomWithQualityWeights(x, design = design, plot = TRUE, 
                         normalize.method = i), edger_norm[1:3], 
  c("quantile", rep("none",2)), SIMPLIFY = FALSE 
)
# create list of normalized data ------
norm_list <- list("Not normalized" = cpm(edger_norm[[1]],normalized.lib.sizes = TRUE), 
                  "EDA_upperQ" = dataOffset, 
                  "TMM" = cpm(edger_norm[[2]],normalized.lib.sizes = TRUE),
                  "TMMwsp" = cpm(edger_norm[[3]],normalized.lib.sizes = TRUE), 
                  "RLE"  = cpm(edger_norm[[4]],normalized.lib.sizes = TRUE),
                  "voom_Quantile" = 2^voomQW_norm[[1]]$E,
                  "voom_TMM" = 2^voomQW_norm[[2]]$E,
                  "voom_TMMwsp" = 2^voomQW_norm[[3]]$E)

pdf(str_glue("{dat_path}/RLE_plots_{todate}_{my_exp}.pdf"))

sapply(seq_along(norm_list) ,function(x, n, i){ 
       EDASeq::plotRLE(x[[i]], col=dgl_filt$colours,
                       outline=FALSE, las=3, 
                       ylab="Relative Log Expression", 
                       cex.axis=1, cex.lab=1, main = str_glue("{n[[i]]}"));
  legend("topright",
       legend= levels(as_factor(dgl_filt$samples$group)),
       fill = levels(as_factor(dgl_filt$colours)),
       bty="n",
       cex = 0.5, inset = c(.01,.01) )}, x = norm_list, n = names(norm_list))
dev.off()

#cpm_TMM <- cpm(x1_TMM,normalized.lib.sizes = TRUE, prior.count=5)

```

## make h-clustering
```{r Hierarchical clustering}
norm_l <- list(
               "unorm" = cpm(edger_norm[[1]],normalized.lib.sizes = TRUE, log=TRUE, prior.count=5),
               "TMM" = cpm(edger_norm[[2]],normalized.lib.sizes = TRUE, log=TRUE, prior.count=5),
               "TMMwsp" = cpm(edger_norm[[3]],normalized.lib.sizes = TRUE, log=TRUE, prior.count=5),
               "voomQ_QW" = voomQW_norm[[1]]$E,
               "voomTMM_QW" = voomQW_norm[[2]]$E,
               "voomTMMwsp_QW" = voomQW_norm[[3]]$E)

hc_methods <- c("ward.D2",
                #"single","median", "centroid","mcquitty")
                "complete",
                "average")
                
l_d <- lapply(norm_l,function(x)dist(t(x)))

l_hc <- sapply(hc_methods, function(x) lapply(l_d, function(y) hclust(y,method = x)))

names(l_hc) <- rep(rownames(l_hc),times = ncol(l_hc))

pdf(str_glue("{dat_path}/hierarchic_clust_{todate}_{my_exp}.pdf"))
for (i in seq_along(l_hc)) {
  myplclust(l_hc[[i]],
            lab.col = dgl_filt$colours, #x1$colours,pal1$value
            #labels = dgl_filt$samples$ ,  
       xlab = str_glue("{names(l_hc[i])}"),
       main = str_glue("{enframe(l_hc[[i]])$value[7]} - {enframe(l_hc[[i]])$value[5]} - group"))
  legend("topright",
       legend = levels(as_factor(dgl_filt$samples$group)),
       fill = levels(as_factor(dgl_filt$colours)), #x1$colours  
       bty="n",
       cex = 0.9)
}
dev.off()
```

## make MDS plot
```{r}
par(mfrow=c(2,2))
par(mar=c(6,5,2,1)+ 0.1)

pdf(str_glue("{dat_path}/MDS_plot_{todate}_{my_exp}.pdf"))

plotMDS(norm_l[[4]],labels = ,
                col=dgl_filt$colours,dim.plot = c(1,2))
legend("topright",
       legend= levels(targets_file$group),
       fill = levels(as_factor(dgl_filt$colours)),
       bty="n",
       cex = 1.5, inset = c(.01,.09) )

plotMDS(norm_l[[4]], labels = ,
                col=dgl_filt$colours,dim.plot = c(3,4),main="MDS plot")

plotMDS(norm_l[[4]], labels = ,
                col=dgl_filt$colours,dim.plot = c(1,3),main="MDS plot")

plotMDS(norm_l[[4]], labels = ,
                col=dgl_filt$colours,dim.plot = c(2,3),main="MDS plot")
plotMDS(norm_l[[4]], labels = ,
                col=dgl_filt$colours,dim.plot = c(2,4),main="MDS plot")
dev.off()
```

## make PCA plot
```{r}
suppressPackageStartupMessages(library(DESeq2))
# make the Deseq dataset -----
dds <- DESeqDataSetFromMatrix(countData = round(dgl_filt$counts),
                              colData = targets_file,
                              design = ~ group )
dds$group <- relevel(dds$group, ref = "Control")

#vst -----
vsd <- vst(dds, blind = FALSE)

head(assay(vsd), 3)

sampleDists <- dist(t(assay(vsd)))
sampleDistMatrix <- as.matrix( sampleDists )
suppressPackageStartupMessages({
library("pheatmap")
library("RColorBrewer")})
sampleDistMatrix <- as.matrix( sampleDists )
rownames(sampleDistMatrix) <- paste( vsd$Run, vsd$group, sep = " - " )
colnames(sampleDistMatrix) <- NULL
colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
pheatmap(sampleDistMatrix,
         clustering_distance_rows = sampleDists,
         clustering_distance_cols = sampleDists,
         col = colors)
#rlog -----
rld <- rlog(dds, blind = FALSE)

dds1 <- estimateSizeFactors(dds)
# check the transformations -----
df <- bind_rows(
  as_tibble(log2(counts(dds1, normalized=TRUE)[, 1:2]+1)) %>%
    mutate(transformation = "log2(x + 1)"),
  as_tibble(assay(vsd)[, 1:2]) %>% mutate(transformation = "vst"),
  as_tibble(assay(rld)[, 1:2]) %>% mutate(transformation = "rlog"))

colnames(df)[1:2] <- c("x", "y")  
# plot them
ggplot(df, aes(x = x, y = y)) + geom_hex(bins = 80) +
  coord_fixed() + facet_grid( . ~ transformation)  
# PCA -----
pdf(str_glue("{dat_path}/PCA_plot_{todate}_{my_exp}.pdf"))
DESeq2::plotPCA(vsd, intgroup = "group")
dev.off()
# MDS -----
mds <- as.data.frame(colData(vsd))  %>%
  cbind(cmdscale(sampleDistMatrix))
ggplot(mds, aes(x = `1`, y = `2`, color = group)) +
  geom_point(size = 3) + coord_fixed()

```
## DE analysis 
```{r}
# DEseq2 ----
dds <- DESeq(dds)
res <- results(dds)
summary(res)
resultsNames(dds)

DEseq2_res_AD_vs_Control <- results(dds, name= "group_AD_vs_Control", alpha = 0.05) %>% 
  as_tibble(rownames = "GeneID") %>% 
  arrange(padj) %>% 
  write_tsv(str_glue("{dat_path}/DEseq2_AD_vs_Ctl_{todate}_{my_exp}.txt"))

hist(DEseq2_res_AD_vs_Control$pvalue[DEseq2_res_AD_vs_Control$baseMean > 1], 
     breaks = 0:20/20,
     col = "grey50", border = "white")
deseq2_p <- DEseq2_res_AD_vs_Control %>% 
  mutate(DEseq2 = if_else(
    padj >= 0.05, 0, if_else(
      log2FoldChange > 0, 1, -1
    )
  )) %>% 
  select(smallRNA = "GeneID", DEseq2 )
# region report for DEseq2 -----
suppressPackageStartupMessages(library(regionReport))
dir.create(str_glue("{dat_path}/DESeq2_report_{todate}"), recursive = TRUE)

report <- DESeq2Report(dds, project = str_glue('DESeq2_{todate}_{my_exp}_HTML_report'),
    intgroup = c('group'), outdir = str_glue("{dat_path}/DESeq2_report_{todate}"),
    output = 'index', theme = theme_bw())


# edgeR  ----
edger_norm %>% names()
plotMD(edger_norm[[4]],column = 3)
plotMD(edger_norm[[4]],column = 4)
x1_TMM <- estimateDisp(edger_norm[[3]], design=design, robust=TRUE)
plotBCV(x1_TMM)
fit <- glmQLFit(x1_TMM, design, robust = TRUE)
head(fit$coefficients)
plotQLDisp(fit)

#con_mat <- makeContrasts(
# IPP_N_vs_INPUT = IPP_N_Ab-IPP_INPUT,
#  levels = design)

edgeR_res_AD_vs_Control <- glmQLFTest(fit,coef = 2)
is.de <- decideTests(edgeR_res_AD_vs_Control)
summary(is.de)

plotMD(edgeR_res_AD_vs_Control,
       status = is.de,
       values = c(1,-1),
       col = c("red","blue"),
       legend="topright")

top <- topTags(edgeR_res_AD_vs_Control,n=100)
res %>% as_tibble()
path <- "data/new_analysis_workflow/test_DE_methods/IP_DATA/"

topedgeR_res_AD_vs_Control <- topTags(edgeR_res_AD_vs_Control, 
                            n = nrow(edgeR_res_AD_vs_Control), 
                            adjust.method = "BH", 
                            sort.by = "PValue", 
                            p.value = 1)$table %>% 
  as_tibble(rownames = "smallRNA") %>% 
  write_tsv(str_glue("{dat_path}/edgeR_res_AD_vs_Ctl_{todate}_{my_exp}.txt"))

hist(topedgeR_res_AD_vs_Control$PValue, 
     breaks = 0:20/20,
     col = "grey50", border = "white")

edgeR_p <- topedgeR_res_AD_vs_Control %>% 
  mutate(edgeR = if_else(
    FDR >= 0.05, 0, if_else(
      logFC > 0, 1, -1
    )
  )) %>% 
  select(smallRNA , edgeR )

# voom  ----
voomQW_norm %>% names
vfit <- lmFit(voomQW_norm[[2]], design=design)
vfit <- contrasts.fit(vfit,coefficients = 2)
vfit <- eBayes(vfit,robust = TRUE,trend = TRUE)

results <- decideTests(vfit)
a <- vennCounts(results)
mfrow.old <- par()$mfrow
par(mfrow=c(1,2))
vennDiagram(a)
vennDiagram(results, 
    include=c("up", "down"),
    counts.col=c("red", "blue"),
    circle.col = c("red", "blue", "green3"))
par(mfrow=mfrow.old)

toplvoom_res_AD_vs_Control  <- topTable(vfit,
                             coef = 1,
                             confint=TRUE,
                             number = nrow(vfit),
                             adjust.method = "fdr",
                             sort.by = "p") %>% 
  as_tibble(rownames = "smallRNA") %>% 
  write_tsv(str_glue("{dat_path}/voomQW_TMM_res_AD_vs_Ctl_{todate}_{my_exp}.txt"))

hist(toplvoom_res_AD_vs_Control$P.Value, 
     breaks = 0:20/20,
     col = "grey50", border = "white")

voomQW_p <- toplvoom_res_AD_vs_Control %>% 
  mutate(voomQW = if_else(
    adj.P.Val >= 0.05, 0, if_else(
      logFC > 0, 1, -1
    )
  )) %>% 
  select(smallRNA , voomQW )

# venn diagram for voom, DEseq2, edger -----
 results <-  deseq2_p %>% 
  inner_join(voomQW_p) %>% 
  inner_join(edgeR_p) %>% select(-smallRNA)

pdf(str_glue("{dat_path}/venn_diagram_DE_{todate}_{my_exp}.pdf"))
vennDiagram(results, 
    include=c("up", "down"),
    counts.col=c("red", "blue"),
    circle.col = c("red", "blue", "green3"))
dev.off()
```


## remove unwanted variation
```{r}
suppressPackageStartupMessages(library('sva'))
# full model
mod <-  model.matrix(~dgl_filt$samples$group+dgl_filt$samples$todate_of_seq,
                     data = dgl_filt$samples)
colnames(mod) <- colnames(mod) %>% 
  str_remove("dgl_filt\\$samples\\$group") %>% 
  str_remove("dgl_filt\\$samples\\$todate_of_seq") 
  str_remove("targets_file\\$Sex") %>% 
  str_remove("targets_file\\$Donors")

# null model
mod0 <- cbind(mod[,1])

# estimate batch and other artifacts
svseq <- svaseq(dgl_filt$counts, mod, mod0, n.sv = 1)
plot(svseq$sv, pch=19, col="blue")

n.sv <- num.sv(voomQW_norm[[1]]$E, mod, method = "leek")
svobj <- sva(voomQW_norm[[1]]$E, mod, mod0,n.sv = n.sv)

modSv <- cbind(mod,svobj$sv)
fit = lmFit(voomQW_norm[[1]],modSv)

# combat
norm_l[["TMM"]]
modcombat <-  model.matrix(~1, data=dgl_filt$samples)
combat_edata <- ComBat(norm_l[["TMM"]], batch = dgl_filt$samples$todate_of_seq,
                       mod = modcombat, par.prior = TRUE,
                       prior.plots = FALSE,
  mean.only = FALSE, ref.batch = NULL, BPPARAM = bpparam("SerialParam"))
pValuesComBat = f.pvalue(combat_edata,mod,mod0)
qValuesComBat = p.adjust(pValuesComBat,method="BH")

#limma remove batch effect
norm_l[["TMM"]] %>% plotMDS(labels = dgl_filt$samples$group, col = dgl_filt$colours)
logCPM <- removeBatchEffect(norm_l[["TMM"]] ,batch = dgl_filt$samples$todate_of_seq)
  plotMDS(logCPM,
  labels = dgl_filt$samples$group,
                col = dgl_filt$colours,dim.plot = c(3,4))

 d <- dist(t(logCPM))
 hc <- hclust(d,method = "ward.D2")
 myplclust(hc,
            lab.col = dgl_filt$colours, #x1$colours,pal1$value
            labels = dgl_filt$samples$group)#pal1$group
      # xlab = str_glue("{names(l_hc[i])}"),
      # main = str_glue("{enframe(l_hc[[i]])$value[7]} - {enframe(l_hc[[i]])$value[5]} - group"))
  legend("topright",
       legend = levels(as_factor(dgl_filt$samples$group)),
       fill = levels(as_factor(dgl_filt$colours)), #x1$colours  
       bty="n",
       cex = 0.9)


# make heatmap -----
suppressPackageStartupMessages(library(ComplexHeatmap))
suppressPackageStartupMessages(library(circlize))

logCPM <- logCPM[str_detect(rownames(logCPM), "piR"),]
filter = apply(logCPM, 1, function(x) length(x[x > 0 ]) > 5)
logCPM_fil <- logCPM[filter, ]
scaled_mat <- t(scale(t(as.matrix(logCPM_fil))))


scaled_mat <- t(scale(t(as.matrix(logCPM))))
hist(scaled_mat)
df <- dgl_filt$samples %>% 
  select(group, todate_of_seq, Sex, Age)

data.frame(Brain_partition = dgl_filt$samples$group,
                 todate_of_seq)  


ha <- HeatmapAnnotation(df = df)
f <- colorRamp2(c(-2,median(scaled_mat),2),c("blue","white","yellow"))

# plot Heatmap
# cluster rows and columns
pdf(str_glue("{dat_path}/Heatmap_{todate}_{my_exp}.pdf"))
Heatmap(scaled_mat,
        top_annotation = ha,
        col=f,
        show_row_dend = TRUE,
        clustering_distance_rows = "spearman",
        clustering_method_rows="ward.D2",
        clustering_distance_columns = "spearman",
        clustering_method_columns ="ward.D2",
        show_row_names = FALSE,show_column_names = F,
        name= "z-score equivalent", 
        column_title = str_glue("Heatmap of {nrow(scaled_mat)} piRNAs"),
        column_title_side="top")  
dev.off()


```


The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
