---
title: "R Notebook, analysis of GSE63501"
output: html_notebook
editor_options: 
  chunk_output_type: console
---
## load libraries
```{r load libraries}
suppressPackageStartupMessages(library('tidyverse'))
suppressPackageStartupMessages(library('data.table'))
```
## add date
```{r date_of_analysis}
date <- gsub(" ","_",format(Sys.time(), "%b %d %Y"))
```
## create the dir for the analysis
```{r}
dir.create("/home/0/SPAR_data_analysis/GSE63501_analysis")
dat_path <- "GSE63501_analysis"
my_exp <- "brain_GSE63501"
```
## import data and make dataframe
```{r create_the_df}
path <- "/home/0/geobrain/GSE63501_16_CTRL_AD"
smallRNA_files <- dir(path, full.names = TRUE,
                      pattern = "smRNA_gene_expression.xls",
                      recursive = TRUE)
# load the list of files in one table -----
DT <- rbindlist(sapply(smallRNA_files,fread,
                       simplify=FALSE,
                       #select = c(1,2,4,6),
                       verbose=getOption("datatable.verbose", TRUE)),
                use.names= TRUE,idcol="file")  %>%
  rename(smallRNA = "#Gene")

# sep "smallRNA" column to multiple
# separate("smallRNA",c("chr","start","end","strand","smallRNA","DQ"), sep = ":") %>%
#  as_tibble()

# make matrix for DE analysis  -----
dt <- DT %>% group_by(file) %>% select(-RPM) %>%
  spread(key = "file", value = "ReadCount" )
# clean the colnames
names(dt)[3:length(names(dt))] <- names(dt)[3:length(names(dt))] %>%
  str_remove(str_glue(".+/GSE63501_16_CTRL_AD/")) %>%
  str_remove(".trimmed_.+")

# load the targets file ----
targ_path <- dir(pattern = "tex.xls", recursive = TRUE)
targ_path <- dir(path = path, 
                 pattern = "Sra", 
                 full.names = T,
                 recursive = TRUE)

targets_file <- fread(targ_path, select = c("Run","classification")) %>%
  as_tibble() %>% mutate(classification = as_factor(classification))

stopifnot(identical(targets_file$Run,names(dt)[3:length(names(dt))]))

# write the results  -----
writting_path <- str_glue("{dat_path}/raw_reads_{date}_{my_exp}_SPAR.txt")
dt %>% write_tsv(writting_path)
```
## Perform various Filtering Methods: EdgeR, NOIseq
```{r}
suppressPackageStartupMessages(library('edgeR'))
suppressPackageStartupMessages(library('NOIseq'))# there is a problem loading it 
# I load it manually 
suppressPackageStartupMessages(library('rafalib'))

# make the matrix  -----
mat <- dt %>%
  select(-GeneClass) %>%
  column_to_rownames("smallRNA") %>%
  as.matrix()

# filtering with NOISEq  -----
noiL <- as.list(1L:3)
names(noiL) <- c("cpm", "Wilcoxon", "Prop")
NOISeq_filt <- lapply(noiL, function(x) NOISeq::filtered.data(mat, 
        factor=targets_file$classification,
        norm=FALSE, 
        method=x, cv.cutoff=100, cpm=1))

# create the design matrix  -----
design <- model.matrix(~targets_file$classification)
colnames(design) <- gsub("targets_file\\$","",colnames(design))
rownames(design) <- targets_file$Run

# create the DGElist ----
targets_file <- targets_file %>% rename(classification = "group")
dgl <- DGEList(counts = mat,
               samples = targets_file,
               lib.size = colSums(mat),
               norm.factors = rep(1,ncol(mat)))

# filterwith EdgeR ----
keep.exprs <- filterByExpr.DGEList(dgl, design = design)
dgl_filt <- dgl[keep.exprs,,keep.lib.sizes=FALSE]

str_glue( "Starting features: {nrow(dgl)}\nedgeR filter, remaining features: {nrow(dgl_filt)}\nCPM approach, remaining features: {nrow(NOISeq_filt[[1]])}\nWilcoxon filter, remaining features: {nrow(NOISeq_filt[[2]])}\nProportion approach filter, remaining features: {nrow(NOISeq_filt[[3]])}")

# if we want to keep noiseq filter instead of edger but use edger object ----
filte_Nois<- dgl[rownames(dgl$counts) %in% rownames(NOISeq_filt[[3]]),,keep.lib.sizes=FALSE]
dgl_NOIs<- DGEList(counts = NOISeq_filt[[3]],
               samples = targets_file,
               lib.size = colSums(NOISeq_filt[[3]]),
               norm.factors = rep(1,ncol(NOISeq_filt[[3]])))
# finally give colours to samples ----
suppressPackageStartupMessages(library('wesanderson'))
pal1 <- wes_palette(length(levels(as_factor(targets_file$group))),
                    name = "Darjeeling1", type = "discrete") %>%
  enframe(name = NULL) %>%
  mutate(
    group = as_factor(levels(targets_file$group)))

# add color to the DGEList 
dgl_NOIs$colours <- inner_join(dgl_NOIs$samples,pal1,by= "group")$value


# plot RLE and other functions from EDASeq
EDASeq::plotRLE(dgl_NOIs$counts, col=dgl_NOIs$colours, outline=FALSE, las=3, ylab="Relative Log Expression", cex.axis=1, cex.lab=1)
EDASeq::plotPCA(dgl_NOIs$counts, col=dgl_NOIs$colours, cex=1, cex.axis=1, cex.lab=1)

dataEDA <- EDASeq::newSeqExpressionSet(counts = dgl_NOIs$counts, phenoData =as.data.frame(targets_file,row.names = targets_file$Run))
EDASeq::meanVarPlot(dataEDA, log=TRUE, ylim=c(0,16))
EDASeq::biasPlot(dataEDA, "gc", log=TRUE, ylim=c(1,5))
```
## proceed with the normalization
```{r}
EDASeq::betweenLaneNormalization()
dataOffset <- EDASeq::betweenLaneNormalization(dataEDA,
                                       which="upper",offset=TRUE)
EDASeq::normCounts(dataOffset)
EDASeq::meanVarPlot(dataOffset, log=TRUE, ylim=c(0,16))
EDASeq::biasPlot(dataOffset, "gc", log=TRUE, ylim=c(1,5))

```



The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
